Drobonin Model Prediction

### 2024-01-29 DrobininModelPredictions.Rmd
Added script for Drobinin model predictions

---
title: "DataSetup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load Structural Data
```{r}
#Load list of included subjects
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

smri_baseline <- smri_1%>% 
  filter(eventname=="baseline_year_1_arm_1")

smri_followup <- smri_1%>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#T1 intensity, might not need
#smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
 # filter(!collection_title=="collection_title") %>% 
 # filter(src_subject_id%in%struc_include$src_subject_id)

#Rename data to fit naming conventions
#????
ogfeatures <- rio::import("~/Documents/FYP/feature_list.txt")

Load Other Data
```{r}
#Puberty
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title")

youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
  filter(!collection_title=="collection_title")

#parent sum scores abcd_mhp01
#youth sum scores abcd_mhy01

#Cognition
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)

#Height/weight metrics
physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)

#Sexual attraction and behavior
gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")

#Grade level, parent report
grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
```

Template for renaming sMRI variables
```{r, eval=F}

@@

#renamed_smri <- subset(renamed_smri, select = -c(269,279))

@@

smri_abcd_model <- renamed_smri_baseline[, 1:180] 
```

Rename sMRI for follow-up
```{r}

renamed_smri_followup <- smri_followup %>% 

#renamed_smri <- subset(renamed_smri, select = -c(269,279))

smri_abcd_model <- renamed_smri_baseline[, 1:180]

@@

Split Samples
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2.Rda")
load("renamed_smri_baseline.Rda")

training_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_1$src_subject_id) %>% 
  select(-c(subjectkey, src_subject_id, sex, eventname))

analysis_sample_baseline <- filter(smri_abcd_model , src_subject_id %in% sample_2$src_subject_id)

save(training_sample_baseline, file="training_sample_baseline.Rds")
save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")

Make dataframe for ML testing
```{r, eval = F}
train_test <- training_sample_baseline %>% 
  #select(1:15) %>% 
  slice_head(n=20)

#validation_test <- df_validation_baseline %>% 
  #select(1:15) %>% 
#  slice_head(n=20)

save(train_test, file="train_test.Rds")
#save(validation_test, file="validation_test.Rds")

Feature list template
```{r, eval=F}
#Template
FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
FS_BrainSeg_Vol
FS_BrainSeg_Vol_No_Vent
FS_BrainSeg_Vol_No_Vent_Surf
FS_LCort_GM_Vol = smri_vol_cdk_totallh,
FS_RCort_GM_Vol = smri_vol_cdk_totalrh,
FS_TotCort_GM_Vol = smri_vol_cdk_total, #total whole brain cortical volume
FS_SubCort_GM_Vol = smri_vol_scs_subcorticalgv,
FS_Total_GM_Vol #maybe just calculate?
FS_SupraTentorial_Vol = smri_vol_scs_suprateialv, #assuming this includes all ventricles
FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
FS_SupraTentorial_No_Vent_Voxel_Count
FS_Mask_Vol
FS_BrainSegVol_eTIV_Ratio
FS_MaskVol_eTIV_Ratio
FS_L_LatVent_Vol = smri_vol_scs_ltventriclelh,
FS_L_InfLatVent_Vol = smri_vol_scs_inflatventlh,
FS_L_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterlh,
FS_L_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexlh,
FS_L_ThalamusProper_Vol = smri_vol_scs_tplh,
FS_L_Caudate_Vol = smri_vol_scs_caudatelh,
FS_L_Putamen_Vol = smri_vol_scs_putamenlh,
FS_L_Pallidum_Vol = smri_vol_scs_pallidumlh,
FS_3rdVent_Vol = 	smri_vol_scs_3rdventricle,
FS_4thVent_Vol = smri_vol_scs_4thventricle,
FS_BrainStem_Vol = smri_vol_scs_bstem,
FS_L_Hippo_Vol = smri_vol_scs_hpuslh,
FS_L_Amygdala_Vol =	smri_vol_scs_amygdalalh,
FS_CSF_Vol = 	smri_vol_scs_csf,
FS_L_AccumbensArea_Vol = 	smri_vol_scs_aal,
FS_L_VentDC_Vol = smri_vol_scs_vedclh,
FS_L_Vessel_Vol
FS_L_ChoroidPlexus_Vol
FS_R_LatVent_Vol = smri_vol_scs_ltventriclerh,
FS_R_InfLatVent_Vol = smri_vol_scs_inflatventrh,
FS_R_Cerebellum_WM_Vol = smri_vol_scs_crbwmatterrh,

# no cerebellum cort vol?
FS_R_Vessel_Vol
FS_R_ChoroidPlexus_Vol
FS_OpticChiasm_Vol

```{r}
#Puberty
#Raw scores, youth
youth_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname== "2_year_follow_up_y_arm_1") %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  select(src_subject_id,interview_age, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected) %>% 
  mutate_at(c(2:5), as.numeric)

#Existing Models
corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")
```

## Format BrainAge
```{r}
corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
```
## Format Puberty Data 
```{r}
#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

youth_pds_followup[youth_pds_followup == 777] <- NA
youth_pds_followup[youth_pds_followup == 999] <- NA

#Youth-report, removes everyone w/o complete data
youth_pds_followup <- youth_pds_followup %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)

youth_pds_use_followup <- youth_pds_followup %>% 
  select(src_subject_id, sex, youth_sum, youth_mean)

#Parent-report, removes everyone w/o complete data
parent_pds_followup[parent_pds_followup == 777] <- NA
parent_pds_followup[parent_pds_followup == 999] <- NA
parent_pds_followup <- parent_pds_followup %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5)

parent_pds_use_followup <- parent_pds_followup %>% 
  select(src_subject_id, sex, parent_sum, parent_mean)
```

## Make big dataframe
```{r}
brainage_pds_followup <- corrected_followup_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_use_followup) %>% 
  left_join(parent_pds_use_followup) %>% 
  left_join(cognition_followup)

#Split sample
smp_size_followup <- floor(0.5 * nrow(brainage_pds_followup))

## set the seed to make your partition reproducible
set.seed(123)
sample_ind_followup <- sample(seq_len(nrow(brainage_pds_followup)), size = smp_size_followup)

sample_1_followup <- brainage_pds_followup[sample_ind_followup, ]
sample_2_followup <- brainage_pds_followup[-sample_ind_followup, ]

#t test for age differences
t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)

#Save samples
save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
save(sample_1_followup, file="Sample1FollowUp.Rda")
save(sample_2_followup, file = "Sample2FollowUp.Rda")

Attach baseline pds/cognition data to sample 2 / abcd model brainage

## Format BrainAge
```{r}
abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id

abcd_brainage_baseline <- abcd_brainage_baseline[, c(6,1:5)]
```

```{r}
abcd_baseline <- sample_2 %>% 
  select(1, 7:16)

abcd_baseline <- left_join(abcd_baseline, abcd_brainage_baseline)

str(abcd_baseline)

save(abcd_baseline, file="abcd_baseline.Rda")
```

#################
Make Dataframe of sMRI features for combined wave model
################
```{r}

>>>>

---
title: "Drobonin Model Prediction"
author: "Lucy Whitmore"
date: "10/6/2023"
output: html_document
---

User notes: this script should be able to make predictions based on the Drobonin et al., 2022 BrainAGE model. The script is currently set up to predict for 2 waves of data, baseline and follow-up (ABCD 2 year followup).

In order to run predictions, you'll need 2 things: the xgboost model, available at https://github.com/GitDro/DevelopmentalBrainAge/blob/master/model/xgboost_9to19_brain_age_mod.rds, and a data frame. The dataframe should have only interview age (in years) and columns for brain measures. IDs can be reattached after predictions are made, the rows won't change in order.

```{r}
library(tidymodels)
library(xgboost) 
```
Code for renaming columns
```{r} 
# The following code should rename ABCD smri variables to match the Drobonin model. Note, not all the features from the model exist in ABCD, so we also have to make empty columns for those.

@@

@@

@@

```{r}
# Import Drobonin BrainAGE Model
mod_1 <- readRDS(
  file = here::here(
    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))

# Note: model1_features_baseline refers to a dataframe with columns only for interview age and brain features, named in accordance with https://github.com/GitDro/DevelopmentalBrainAge/blob/master/feature_list.txt

# check for any columns that aren't in model
names.use <- names(model1_features_baseline)[!(names(model1_features_baseline) %in% mod_1$fit$feature_names)]

# baseline
# remove those columns
model1_features_baseline_clean <- model1_features_baseline %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_baseline = as.matrix(model1_features_baseline_clean)
newdata_baseline = newdata_baseline[ , mod_1$fit$feature_names]

### Follow-Up
# remove those columns
model1_features_followup_clean <- model1_features_followup %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_followup = as.matrix(model1_features_followup_clean)
newdata_followup = newdata_followup[ , mod_1$fit$feature_names]
```

run predictions - BASELINE
```{r}

# predict brainage - baseline

# Predict
# This prediction method gives the output as a vector, we'll fix that later on
brainage_mod1_baseline <-
  mod_1$fit %>%
  predict(newdata = newdata_baseline)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_baseline <- as.data.frame(brainage_mod1_baseline) %>%
  rename(.pred = brainage_mod1_baseline) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_baseline$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# bias correction
# Slope and intercept are hardcoded
bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
bias_slope <- 0.55 # xgb_bias_mod$coefficients[["truth"]]

# create bias corrected data frame
brain_age_df_mod1_baseline_corrected <- brain_age_df_mod1_baseline %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_baseline_corrected$src_subject_id <- harmonized_mri_baseline$src_subject_id
```

run predictions - FOLLOW-UP
```{r}
# predict brainage - followup
brainage_mod1_followup <-
  mod_1$fit %>%
  predict(newdata = newdata_followup)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_followup <- as.data.frame(brainage_mod1_followup) %>%
  rename(.pred = brainage_mod1_followup) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_followup$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# create bias corrected data frame
brain_age_df_mod1_followup_corrected <- brain_age_df_mod1_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_followup_corrected$src_subject_id <- harmonized_mri_followup$src_subject_id
```
